Online {\IterComp}

### 2017-09-12 abstract.tex
Abstract

\begin{abstract}
Until recently, most of the existing work in {\itercomp} had been focusing on
finding the best optimisation through repeated runs using a single input.
However, in real-world online scenarios, the user rarely executes a program with
the same input multiple times and most of the programs are complex enough that a
single input case does not capture the whole range of possible scenarios and
program behaviours.

In this thesis, we focus on enabling {\itercomp} in \textit{online} scenarios.
Because of the restriction on repeating the execution with same input, it is
not always feasible to measure the real speedup when comparing optimisations.
To this end, we propose a \textit{work efficiency} metric to guide
online {\itercomp}, which is able to achieve about 80\% of the performance
of an offline oracle, with improvements of up to 20\%.

Moreover, because we need to instrument the program for measuring work
efficiency, providing a low-overhead profiling is essential.
In order to reduce the profiling overhead, we also propose two relaxation
algorithms which provide a trade-off between measurement accuracy and overhead.
Our results show that the two relaxation algorithms can reduce the average
overhead by 43\% and $2.1\times$ over the optimal work profiling, observing
up to about $5\times$ of overhead reduction.
\end{abstract}

>>>>

\begin{abstract}

    Iterative or adaptive compilation can dramatically improve the performance of a program by searching for the best optimisation
    settings. However, it relies upon repeatedly evaluating and comparing optimisations off-line on representative inputs. If the
    developers fail to guess what is a representative input or usage patterns change over time, then the wrong optimisations may be
    selected. Until this paper, it has been impossible to search online over actual user inputs. Program side effects, interactivity and
    difficulties capturing the system state prevent repeated execution. Since each input may be executed only once and each input may
    entail a different amount of work, the search cannot use runtime to compare optimisations. Unless we find a way to compare
    optimisations even when inputs are executed only once, we will be unable to use real user input for online iterative compilation, users
    will not have the best optimised programs, and performance and energy efficiency will suffer.

    This paper demonstrates online iterative compilation. We compare optimisations by work efficiency, or work divided by time, executing
    each input only once. We develop an instrumented, low overhead, work efficiency metric with optimal probe placement. Then we further
    reduce the overhead by removing high-cost probes whose absence introduces only small errors to the work metric, without significantly
    affecting the outcome of the iterative compilation. We give two variants for probe removal optimisation: the first offers strong
    guarantees about the maximum possible error, while the second targets the average, whole program error.

    Our online method yields programs that get 80\% of the speedup an offline oracle achieves but without executing any input more than
    once, which the oracle cannot do. Through probe removal we bring the instrumentation overhead down to only 4\% on average, making our
    approach suitable for regular use. We show practical online iterative compilation for the first time, optimising programs according to
    the range of inputs users actually present.

\end{abstract}
